### 2018-08-07 paper/intro.tex
typo in intro, started conclusion edits, more needed

This thirst for seamless sharing explains the rise and ongoing popularity
of personal sharing platforms such as Dropbox, Google Docs,
OneDrive, Amazon Drive, etc. These solutions allow users to share data between any device, with anyone, almost
anytime, anywhere. They achieve this feat by relying heavily on
\emph{remote cloud services}, which have now become a key enabler of ubiquitous ecosystems. 
% argument: cloud -> danger, 
%% From a technical perspective, one major trend to provide a seamless
%% sharing of digital contents among users' devices is to leverage on the
%% cloud, which acts as a key enabler for ubiquitous computing. Users' data
%% are pushed online, into the cloud with either Dropbox, Google Docs,
%% OneDrive, Amazon Drive, etc. to make it accessible anytime,
%% anywhere to the users' devices.
%% \commentFT{Removing the following, as seems to be a detour from main argument: \emph{In the last decades, a lot of research work has
%% emerged to promote the use of this offloading strategy to overcome scarce
%% resources of users' devices, and to provide an always-on availability
%% \cite{Zhang:2014,Gordon:2012,Chun:2011}. It paves the way of different kind
%% of clouds, such as mobile cloud computing, mobile-edge computing
%% and/or fog computing, that aim to use the nodes from the cloud
%% infrastructure that are the closest to end users
%% for offloading with an improved quality of service and user experience.}}
%% Hence, the cloud, and its variants constitute a perfect complement
%% to perform storage and computation on behalf of the resource
%% constrained devices.
Unfortunately, this reliance on the cloud comes at a strong price: 
%% users and their devices become totally tethered to the cloud ecosystem and its providers. As
%% a direct consequence,
users lose control over their own data, which is distributed over
multiple data centers, often in foreign jurisdictions, making it difficult for most users to grasp, review, and assess the privacy, and
confidentiality guaranties they can expect. In practice, user data hosted in the cloud is routinely harvested, often without explicit
consent, either by cloud providers or third parties.

Although data encryption is often proposed as a first line of defense for cloud-hosted data, it is unfortunately  insufficient~\cite{granick_2017, HooffLZZ15, HarnikPS10}. Metadata such as a user'
location and activity can still be tracked, including access
control histories, thus revealing what each user shared with whom and when. 
%  because it does not protect meta-data information
The exposure of metadata seriously weakens user privacy, as famously emphasized by a former NSA and CIA director: "\emph{We kill people based on metadata.}''~\cite{NaughtonTheGuardian2016} 
In this model, user devices, masquerading as allies, get turned into %% tracking and/or monitoring \squad{}s, and serve as
highly efficient spies over their owners.

Constructing a system that provides fully anonymous content sharing while avoiding the above pitfalls is unfortunately highly
challenging. The last two decades have witnessed many attempts at building privacy-preserving and anonymous P2P data-sharing
networks~\cite{Clarke:2001,Gnunet2002,Freedman:2002,Nambiar:2006,Rennhard:2002}.
These solutions typically leverage \emph{onion
 routing}~\cite{Chaum:1981}, with some improvements to perform node
 discovery in a distributed manner, a step that is otherwise
centralized in traditional onion routing schemes. However, because of the high
churn typically experienced by P2P networks, the performance of these systems has been
discouraging~\cite{LeBlond:2013}; churn causes routes to disappear and be
reconstructed frequently, which is particularly costly. By contrast, Tor
\cite{Tor}, one of the most popular anonymous networks,
trades lower anonymity guaranties for a much better performance. Tor's lower protection arises from
a partially centralized design that hinges on a set of \emph{directory
servers}, which are used to advertise trusted nodes that are available to bootstrap onion
routes.

In this paper, we demonstrate how good performance can be reconciled
with strong privacy protection in data-sharing systems. We present
\name{}, a novel fully decentralized, autonomous,
% approach that provides a 
and self-organizing file sharing service between the \squad{}s of different users that preserves user privacy
in terms of both data \emph{and} metadata. \commentFT{Moving to 3 ideas, to highlight proxemics in its own right} Our proposal hinges on three
key ideas: \emph{Firstly}, the emergence of \emph{\squad{}s}, the set of connected devices owned and used by one user, 
%% the recent advances of smartphones and other
%% mobile devices enable e-squads and it has thus
%% become possible to
makes it possible to apply P2P technology to one's own devices
alone, thereby reducing one's exposure to security and privacy threats
from third parties. \emph{Secondly}, we consider \emph{proxemics relationships} as a key enabler to initiate file
transfers. Physical proximity allow users to share
routing information out-of-band.
%, e.g. from their \squad. \commentFT{I did not get what 'from their \squad' mean.} 
In contrast to state-of-the-art P2P systems, users of \name agree out-of-band to share a file, but
do not keep it available for anyone else in the future. \emph{Thirdly}, we
exploit distributed machine learning techniques to predict the future
availability of user devices, and use this ability to implement a
\emph{probabilistic predictive onion routing} mechanism. The
predictive nature of our routing, exploited in a stateless way, allows us to overcome the inherent fragility of pure P2P data-sharing schemes.

@@

\sonja{commented out my changes, left Daniel's in merge conflict, left
  my changes on paper outline} 
Our contributions are as follows: 
\begin{itemize}

@@

  \item We provide the \ac{SPOR} node selection strategy which, given a secure 
    \ac{RPS} protocol and availability estimates, selects nodes to populate the 
    layers of Sphinxes to maximize the availability of the route --- this 
    yields reliable routes even with low-availability nodes.

  \item We provide a protocol for a device swarm to estimate the availabilities 
    of its own devices.
    These availabilities can then be used with \ac{SPOR}.
    \commentDaniel{We should add some more details, maybe.}

  \item We use \ac{SPOR} to design \name: a novel, fully decentralized, 
    autonomous and self-organizing file transfer service among users.

@@

In the following, in Section~\ref{system-model}, we introduce the
system and adversary models as well as the desired security and privacy properties of \name. Section~\ref{design}
introduces the design of our proposed solution. In
Sections~\ref{security-discussion} and \ref{Performance},
respectively, we evaluate the security and performance of our
approach. We set \name into context with related work in
Section~\ref{RelatedWork} and conclude with Section~\ref{conclusion}.

>>>>

This thirst for seamless sharing explains the rise and ongoing popularity
of personal sharing platforms such as Dropbox, Google Docs,
OneDrive, Amazon Drive, etc. These solutions allow users to share data between any device, with anyone, almost
anytime, anywhere. They achieve this feat by relying heavily on
\emph{remote cloud services}, which have now become a key enabler of ubiquitous ecosystems. 
% argument: cloud -> danger, 
%% From a technical perspective, one major trend to provide a seamless
%% sharing of digital contents among users' devices is to leverage on the
%% cloud, which acts as a key enabler for ubiquitous computing. Users' data
%% are pushed online, into the cloud with either Dropbox, Google Docs,
%% OneDrive, Amazon Drive, etc. to make it accessible anytime,
%% anywhere to the users' devices.
%% \commentFT{Removing the following, as seems to be a detour from main argument: \emph{In the last decades, a lot of research work has
%% emerged to promote the use of this offloading strategy to overcome scarce
%% resources of users' devices, and to provide an always-on availability
%% \cite{Zhang:2014,Gordon:2012,Chun:2011}. It paves the way of different kind
%% of clouds, such as mobile cloud computing, mobile-edge computing
%% and/or fog computing, that aim to use the nodes from the cloud
%% infrastructure that are the closest to end users
%% for offloading with an improved quality of service and user experience.}}
%% Hence, the cloud, and its variants constitute a perfect complement
%% to perform storage and computation on behalf of the resource
%% constrained devices.
Unfortunately, this reliance on the cloud comes at a strong price: 
%% users and their devices become totally tethered to the cloud ecosystem and its providers. As
%% a direct consequence,
users lose control over their own data, which is distributed over
multiple data centers, often in foreign jurisdictions, making it difficult for most users to grasp, review, and assess the privacy, and
confidentiality guarantees they can expect. In practice, user data hosted in the cloud is routinely harvested, often without explicit
consent, either by cloud providers or third parties, only partially
mitigated by legal measures such as the GDPR.

Although data encryption is often proposed as a first line of defense for cloud-hosted data, it is unfortunately  insufficient~\cite{granick_2017, HooffLZZ15, HarnikPS10}. Metadata such as a user'
location and activity can still be tracked, including access
control histories, thus revealing what each user shared with whom and when. 
%  because it does not protect meta-data information
The exposure of metadata seriously weakens user privacy, as famously emphasized by a former NSA and CIA director: "\emph{We kill people based on metadata.}''~\cite{NaughtonTheGuardian2016} 
In this model, user devices, masquerading as allies, get turned into %% tracking and/or monitoring \squad{}s, and serve as
highly efficient instruments spying on their owners.

Constructing a system that provides fully anonymous content sharing while avoiding the above pitfalls is unfortunately highly
challenging. The last two decades have witnessed many attempts at building privacy-preserving and anonymous P2P data-sharing
networks~\cite{Clarke:2001,Gnunet2002,Freedman:2002,Nambiar:2006,Rennhard:2002}.
These solutions typically leverage \emph{onion
 routing}~\cite{Chaum:1981}, with some improvements to perform node
 discovery in a distributed manner, a step that is otherwise
centralized in traditional onion routing schemes. However, because of the high
churn typically experienced by P2P networks, the performance of these systems has been
discouraging~\cite{LeBlond:2013}; churn causes routes to disappear and be
reconstructed frequently, which is particularly costly. By contrast, Tor
\cite{Tor}, one of the most popular anonymous networks,
trades lower anonymity guarantees for a much better
performance. \sonja{lower than who?} Tor's lower protection arises from
a partially centralized design that hinges on a set of \emph{directory
servers}, which are used to advertise trusted nodes that are available to bootstrap onion
routes.

In this paper, we demonstrate how good performance can be reconciled
with strong privacy protection in file transfer systems. We present
\name{}, a novel fully decentralized, autonomous,
% approach that provides a 
and self-organizing file transfer service between the \squad{}s of different users that preserves user privacy
in terms of both data \emph{and} metadata. \commentFT{Moving to 3 ideas, to highlight proxemics in its own right} Our proposal hinges on three
key ideas: \emph{Firstly}, the emergence of \emph{\squad{}s}, the set of connected devices owned and used by one user, 
%% the recent advances of smartphones and other
%% mobile devices enable e-squads and it has thus
%% become possible to
makes it possible to apply P2P technology to one's own devices
alone, thereby reducing one's exposure to security and privacy threats
from third parties. \emph{Secondly}, we consider \emph{proxemics relationships} as a key enabler to initiate file
transfers. Physical proximity allow users to share
routing information out-of-band.
%, e.g. from their \squad. \commentFT{I did not get what 'from their \squad' mean.} 
In contrast to state-of-the-art P2P systems, users of \name agree out-of-band to share a file, but
do not keep it available for anyone else in the future. \emph{Thirdly}, we
exploit distributed machine learning techniques to predict the future
availability of user devices, and use this ability to implement a
\emph{probabilistic predictive onion routing} mechanism. The
predictive nature of our routing, exploited in a stateless way, allows
us to overcome the inherent fragility of pure P2P file transfer schemes.

@@

Our contributions are as follows: 
\begin{itemize}

@@

  \item We provide the \ac{SPOR} node selection strategy for routing which, given a secure 
    \ac{RPS} protocol and availability estimates, selects nodes to populate the 
    layers of Sphinxes to maximize the availability of the route --- this 
    yields reliable routes even with low-availability nodes.

  \item We provide a protocol for a device squad to estimate the availabilities 
    of its own devices, based on Markov models and \squad overlays.
    These availabilities can then be used with \ac{SPOR}.
   % \commentDaniel{We should add some more details, maybe.}

  \item We use \ac{SPOR} to design \name \sonja{removed for
      redundancy, rephrased: a novel, fully decentralized, 
    autonomous and self-organizing} as a privacy-preserving file transfer service among
  users without third-party storage.

@@

In the following, in Section~\ref{system-model}, we state the
system and adversary models as well as the desired security and privacy properties of \name. Section~\ref{design}
introduces the design of our proposed solution. In
Sections~\ref{security-discussion} and \ref{Performance},
respectively, we evaluate the security and performance of our
approach. We set \name into context with related work in
Section~\ref{RelatedWork} and conclude with Section~\ref{conclusion}.

### 2018-08-07 paper/security-discussion.tex

The second change is that we use several nodes per layer as opposed to only 
one.
The proofs of Sphinx rest on the results of \textcite{CLOnionRouting}, which 
are results for ordinary \ac{OR} with only one node per layer.

Analogously to traditional \ac{OR},  we also reveal the previous and next sets 
of nodes to the adversary. 
We note that this is actually quite natural as a malicious adversary can drop 
any incoming traffic to a router and wait for the timeout and
retransmission, \ie when the 
node has selected a new node from the next layer and retried sending.
As such, the adversary can learn all nodes in the layer --- just as for 
traditional \ac{OR}, regardless of whether the set making up a layer contains a
single or several nodes.

>>>>

The second change is that we use several nodes per layer as opposed to only 
one.
The proofs of Sphinx rest on the results of \textcite{CLOnionRouting}, which 
are results for traditional \ac{OR} with only one node per layer.
We must thus ensure that the proofs of \citeauthor{CLOnionRouting} also holds 
for sets of nodes.

Analogously to traditional \ac{OR}, we also reveal the previous and next sets 
of nodes to the adversary.
We note that this is actually quite natural as a malicious adversary can drop 
any incoming traffic to a router and wait for the retransmission, \ie when the 
node has selected a new node from the next layer and retried sending.
As such, the adversary can learn all nodes in the layer --- just as for 
traditional \ac{OR}, regardless of whether the set making up a layer contains a
single or several nodes.

### 2018-08-07 paper/sphinxes.tex
fixing conflict resolution error
fixing compil error

\subsection{Sphinxes: the Sphinx header format Extended for Sets of nodes}%
\label{Sphinxes}

\Textcite{Sphinx} designed \iac{UC} secure message format for mixes, called 
Sphinx.
Since it is \ac{UC} secure, it provably implements an ideal onion routing 
functionality.
\Ie, it provides bitwise unlinkability on the forward route, hides the path 
length, makes replies indistinguishable from forward messages and resists 
tagging attacks.
It is also compact with only 224 bytes of header for a five-hop path with 
128-bit security. \sonja{what about the CL properties?}
Sphinx only provides one node per onion layer; in this section we extend Sphinx 
to Sphinxes, which allows for multiple nodes per onion layer.
(See \cref{fig:OR-POR}.)
% AL: Bounded random-walk....
As such, Sphinxes could be described as a bounded random-walk protocol, \ie the 
each node selects the next node from a set restricted by the source.

\begin{figure}
  \centering
  %  \includegraphics[width=\linewidth]{figures/OR-POR.png}
    \includegraphics[width=0.65\linewidth]{figures/principle_SPOR.pdf}
  \caption{\label{fig:OR-POR}%
    Illustration of the conceptual difference between traditional onion routing

    and mixing (OR, top) and \ac{SPOR} (bottom).

    %\commentDaniel{This figure should be fixed, issue \#42.}
  }
\end{figure}

@@

We will now give an overview of the Sphinx header format.
The idea is not to give a detailed description of Sphinx, but to give the 
reader enough understanding of Sphinx to understand the changes we make to 
achieve Sphinxes.
For a more detailed understanding of Sphinx, we refer the reader to the 
excellent description in the original Sphinx paper~\cite{Sphinx}.

@@

\subsubsection{The Sphinxes extension for sets of nodes}

@@

\subsubsection{Sphinxes algorithms and \acf*{SPOR}}

Sphinxes provides the following algorithms:
\begin{itemize}
  \item \(\CreateHeader\) creates a header.

@@

If \(L' = \{\rdvnode\}\), the packet payload contains another Sphinxes packet 
to be forwarded.
Decrypt the payload \(\delta\) to get the mix packet \((M', \delta')\gets 
  \prp^{-1}(\hash_\prp(s_i), \delta)\).
Run the \(\ProcessHeader\) algorithm with \((M', \delta')\) and 
\(\ProcessMessage\) as input.

>>>>

\subsection{\Sphinxes: the Sphinx header format Extended for Sets of nodes}%
\label{Sphinxes}

\Textcite{Sphinx} designed \iac{UC} secure message format for mixes, called 
Sphinx.
Since it is \ac{UC} secure, it provably implements an ideal onion routing 
functionality.
\Ie, it provides bitwise unlinkability on the forward route, hides the path 
length, makes replies indistinguishable from forward messages and resists 
tagging attacks.
It is also compact with only 224 bytes of header for a five-hop path with 
128-bit security. \sonja{what about the CL properties?}
Sphinx only provides one node per onion layer; in this section we extend Sphinx 
to \Sphinxes, which allows for multiple nodes per onion layer.
(See \cref{fig:OR-POR}.)
% AL: Bounded random-walk....
As such, \Sphinxes could be described as a bounded random-walk protocol, \ie the 
each node selects the next node from a set restricted by the source.

\begin{figure}
  \centering
  %  \includegraphics[width=\linewidth]{figures/OR-POR.png}
    \includegraphics[width=0.65\linewidth]{figures/principle_SPOR.pdf}
  \caption{\label{fig:OR-POR}%
    Illustration of the conceptual difference between traditional onion routing 
and mixing (OR, top) and \ac{SPOR} (bottom).
    %\commentDaniel{This figure should be fixed, issue \#42.}
  }
\end{figure}

@@

We will now give an overview of the Sphinx header format.
The idea is not to give a detailed description of Sphinx, but to give the 
reader enough understanding of Sphinx to understand the changes we make to 
achieve \Sphinxes.
For a more detailed understanding of Sphinx, we refer the reader to the 
excellent description in the original Sphinx paper~\cite{Sphinx}.

@@

\subsubsection{The \Sphinxes extension for sets of nodes}

@@

\subsubsection{\Sphinxes algorithms and \acf*{SPOR}}

\Sphinxes provides the following algorithms:
\begin{itemize}
  \item \(\CreateHeader\) creates a header.

@@

If \(L' = \{\rdvnode\}\), the packet payload contains another \Sphinxes packet 
to be forwarded.
Decrypt the payload \(\delta\) to get the mix packet \((M', \delta')\gets 
  \prp^{-1}(\hash_\prp(s_i), \delta)\).
Run the \(\ProcessHeader\) algorithm with \((M', \delta')\) and 
\(\ProcessMessage\) as input.

### 2018-08-07 paper/squad-overlay-design.tex

The exchanged messages form a sequence, ordered by timestamp, populated by three message types:
\begin{itemize}
	\item \texttt{CONNECTED<$d, t_{step}$>}: Sent by device $d$ when it is used by its user at time step\footnote{As far as devices connections are concerned, time is discretized: we consider that each interaction lasts the same time.} $t_{step}$. $d$ is otherwise considered offline at this time. Devices use this information to predict their probability of remaining online, as presented in section~\ref{ssec:device_availability};
	\item \texttt{SENDING<\sendingdevice, \fileid>}: When the user decides to send a file $f$ located on a device \sendingdevice, \sendingdevice informs its peers of this fact, and joins the file ID \fileid. Squad members will use this information to forward acknowledgments from the remote to \sendingdevice;
	\item \texttt{RECEIVING<\recdevice, \fileid>}: Reciprocally, when the user starts receiving a file $f$ onto a device \recdevice, \recdevice informs its peers of this fact, and joins the file ID \fileid. Squad members will use this information to forward the file's content from the remote to \recdevice.
\end{itemize}

@@

% Every time Bob successfully receives a chunk $\chunk_i$, \recdevice sets the bit $\bitfield[i]$, 
% To acknowledge the chunk's reception, \recdevice sends \bitfield back to Alice, through a unique route $\mathcal{H}'_{\chunk_i}$.

>>>>

The exchanged messages form a sequence, ordered by timestamp, populated by three message types:
\begin{itemize}
	\item \texttt{CONNECTED<$d, t_{step}$>}: Sent by device $d$ when it is used by its user at time step\footnote{As far as devices connections are concerned, time is discretized: we consider that each interaction lasts the same time.} $t_{step}$. $d$ is otherwise considered offline at this time. Devices use this information to predict their probability of remaining online, as presented in section~\ref{ssec:device_availability};
	\item \texttt{SENDING<\sendingdevice, \fileid>}: When the user decides to send a file $f$ located on a device \sendingdevice, \sendingdevice informs its peers of this fact, and joins the file ID \fileid. \Squad members will use this information to forward acknowledgments from the recipient to \sendingdevice;
	\item \texttt{RECEIVING<\recdevice, \fileid>}: Reciprocally, when the user starts receiving a file $f$ onto a device \recdevice, \recdevice informs its peers of this fact, and joins the file ID \fileid. \Squad members will use this information to forward the file's content from the sender to \recdevice.
\end{itemize}

@@

### 2018-08-07 paper/system-model.tex

Friendly surveillance can be mitigated by ensuring security properties 
typically associated with onion routing, as defined by Camenisch and Lysyanskaya~\cite{CLOnionRouting}: 
% We combine the requirements from
% the \ac{P2P} setting with those from onion routing for anonymity.
% Camenisch and Lysianskaya \cite{CLOnionRouting} define security properties 
% for onion routing:
\textit{onion-correctness, integrity}, and \textit{security}. %(wrap resistance,
                                %indistinguishability)
Informally, \textit{onion-correctness} means that if an onion is formed
correctly, processed by the correct routers in the correct order, then
the correct message is received by the last router. \textit{Onion-integrity}
means that even for an onion created by an adversary, the path length
is bounded by $N$. \textit{Onion-security} means that an adversary sending a
challenge onion \david{what is a challenge onion?} to a router and able to see the outcome of how the
onion is processed, cannot distinguish between whether that onion
corresponds to a particular message and route or not. This is also
true for a \emph{re-wrapped} onion, \ie a challenge onion wrapped in
another layer. The formal definitions are in~\cite{CLOnionRouting} along
with a theorem that states that an onion routing scheme satisfying
these three properties, combined with secure point-to-point channels,
is UC-secure \david{what does it means, it has never been introduced
  before, what do you mean?}.
These properties cover the concerns about
revealing to others that Alice shares a file with Bob in the first
place. In \name, we additionally derive the property of
\emph{device-unlinkability} from the \ac{P2P} requirements, meaning that
using \name does not reveal the \squad a device belongs to nor its
owner. Thus, no information on the user's behavior (e.g. being online)
can be infered through \name. \sonja{insert more formal definition
  here, and perhaps explanation that we need it because we introduce
  availability information}

>>>>

Friendly surveillance can be mitigated by ensuring security properties 
typically associated with onion routing, as defined by Camenisch and Lysyanskaya~\cite{CLOnionRouting}: 
% We combine the requirements from
% the \ac{P2P} setting with those from onion routing for anonymity.
% Camenisch and Lysianskaya \cite{CLOnionRouting} define security properties 
% for onion routing:
\textit{onion-correctness, integrity}, and \textit{security}. %(wrap resistance,
                                %indistinguishability)
Informally, \textit{onion-correctness} means that if an onion is formed
correctly, processed by the correct routers in the correct order, then
the correct message is received by the last router. \textit{Onion-integrity}
means that even for an onion created by an adversary, the path length
is bounded by $N$. \textit{Onion-security} means that an adversary sending an
 onion to a router as a challenge
 and able to see the response (the outcome of how the
onion is processed), cannot distinguish between whether that onion
corresponds to a particular message and route or not. This is also
true for a \emph{re-wrapped} onion, \ie a challenge (onion) wrapped in
another layer. The formal definitions are in~\cite{CLOnionRouting} along
with a theorem that states that an onion routing scheme satisfying
these three properties, combined with secure point-to-point channels,
is universal composability (UC)-secure, \ie the protocol remains
secure even if arbitrarily composed with other instances of the same or other protocols.
These properties cover the concerns about
revealing to others that Alice shares a file with Bob in the first
place. In \name, we additionally derive the property of
\emph{device-unlinkability} from the \ac{P2P} requirements above, meaning that
using \name does not reveal the \squad a device belongs to nor its
owner. Thus, no information on the user's behavior (e.g. being online)
can be infered through \name. 
\sonja{insert more formal definition
  here, and perhaps explanation that we need it because we introduce
  availability information}
